---
title: "p8105_hw1_ei2262"
output: github_document
---

```{r, echo=FALSE, include = FALSE}
library(tidyverse)
```

# Problem 1

## Description of `penguins` Dataset

```{r, echo = FALSE}
data("penguins", package = "palmerpenguins")
#Viewing dataset (name/values)
penguins
```
The dataset `penguins` focuses on the following descriptive characteristics of penguins:

1. Type of species (Adélie, Chinstrap and Gentoo)
2. Island in Palmer Archipelago, Antarctica (Biscoe, Dream or Torgersen)
3. Bill length of penguin (in mm)
4. Bill depth  of penguin (in mm)
5. Flipper length of penguin (in mm)
6. Body mass of penguin (in g)
7. Sex of penguin (female or male)
8. Study year (2007, 2008, or 2009)

The dataset contains `r nrow(penguins)` rows and `r ncol(penguins)` columns. 
The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE),2)` mm.

## Scatterplot

The following is a scatterplot with `bill_length_mm` on the x-axis and
`flipper_length_mm`on the y-axis. The dataset is color coordinated based on the 
variable `species`.
```{r, echo = FALSE}
ggplot(data = penguins, aes(x=bill_length_mm,y=flipper_length_mm))+geom_point(aes(color=species))
```

The scatterplot was exported as a pdf to my project directory using `ggsave`. It was named "penguinsplot." 
```{r, echo = FALSE, include = FALSE}
ggsave("penguinsplot.pdf", units = "mm")
```

# Problem 2

## Creating a dataframe called `hw1_df`
```{r, collapse = TRUE}
hw1_df = tibble(
  vec_numeric = rnorm(n = 10),
  vec_logical = vec_numeric > 0,
  vec_character = c("d","s","h","o","m","e","w","o","r","k"),
  vec_factor = factor(c("high", "med","high","high","high","low","low","low","med","med"))
)

mean(pull(hw1_df, vec_numeric))

mean(pull(hw1_df, vec_logical))

mean(pull(hw1_df, vec_character))

mean(pull(hw1_df, vec_factor))
```
After creating the dataframe, I tried calculating the mean of each variable. I was able to calculate the mean for two variables: *vec_numeric* and *vec_logical*. When I calculated the mean for variables *vec_character* and *vec_factor*, I got an output of **NA**.

## Applying `as.numeric`
```{r, eval = FALSE, include = TRUE}
logical_num = as.numeric(pull(hw1_df, vec_logical))
mean(logical_num)

character_num = as.numeric(pull(hw1_df, vec_character))
mean(character_num)

factor_num = as.numeric(pull(hw1_df, vec_factor))
mean(factor_num)
```
#### What happens?
After applying the `as.numeric` function to the logical, character, and factor variables, I was able to calculate the mean of the logical and factor variables. The character variable produces **NA** as the output.   

#### Why?
A logical vector has two possible levels, true or false. In `hw1_df`, a logical operator (">") is used to identify the inputs (elements of the random sample) and return a logical value. Therefore, when `as.numeric` is applied, the observations in a logical vector is converted to a 0 (false) or to a 1 (true). 0 and 1 are both numeric values and can be used to calculate a mean value. In the logical vector I created in `hw_1df`, 

A factor vector can have multiple levels. The one created in `hw1_df` has 3 levels: high, medium, and low. When `as.numeric` was applied to it, each level was converted to a corresponding number, such as 1 for high, 2 for medium, and 3 for low. Numbers 1 - 3 were then used to calculate a mean value.  

A character vector is not numeric or logical because it is made up of characters. In the character vector created in `hw1_df`, some of values in the character strings are not properly formatted. Therefore, `as.numeric` cannot convert the characters to numeric values. This results in the **NA** output with the warning message: NAs introduced by coercion.

##### Understanding the output of `mean()`

Analyzing the results of `mean()` after applying `as.numeric` helped me understand what happens when I try to take the mean of the variables in my dataframe.

