---
title: "p8105_hw1_ei2262"
output: github_document
---

```{r, echo=FALSE, include = FALSE}
library(tidyverse)
```

# Problem 1

## Description of `penguins` Dataset
```{r, echo = FALSE}
data("penguins", package = "palmerpenguins")
#Viewing dataset (name/values)
penguins
```
The dataset `penguins` focuses on the following descriptive characteristics of penguins:

1. Type of species
2. Island of origin
3. Bill length of penguin(in mm)
4. Bill depth  of penguin(in mm)
5. Flipper length of penguin(in mm)
6. Body mass of penguin(in g)
7. Sex of penguin
8. Year penguin was born

```{r, echo = FALSE}
#Number of rows 
nrow(penguins)
#Number of columns
ncol(penguins)

#Mean flipper length
mean(penguins$flipper_length_mm, na.rm = TRUE)
```

The dataset contains 344 rows and 8 columns. 
The mean flipper length is 200.92 mm.

## Scatterplot

The following is a scatterplot with `bill_length_mm` on the x-axis and
`flipper_length_mm`on the y-axis. The dataset is color coordinated based on the 
variable `species`.
```{r, echo = FALSE}
ggplot(data = penguins, aes(x=bill_length_mm,y=flipper_length_mm))+geom_point(aes(color=species))
```

The scatterplot was exported as a pdf to my project directory using `ggsave`. It was named "penguinsplot." 
```{r, echo = FALSE, include = FALSE}
ggsave("penguinsplot.pdf", units = "mm")
```

# Problem 2
```{r}
hw_df = tibble(
  vec_numeric = rnorm(n = 10),
  vec_logical = vec_numeric > 0,
  vec_character = c("d","s","h","o","m","e","w","o","r","k"),
  vec_factor = factor(c("high", "med","high","high","high","low","low","low","med","med"))
)

mean(pull(hw_df, vec_numeric))
mean(pull(hw_df, vec_logical))
mean(pull(hw_df, vec_character))
mean(pull(hw_df, vec_factor))
```
After creating the dataframe, I tried calculating the mean of each variable. The variables that produced an output
were *vec_numeric* and *vec_logical*. The variables that did not produce an output wer *vec_character* and *vec_factor*.

# Applying `as.numeric`
```{r, eval = FALSE, include = TRUE}
logical_num = as.numeric(pull(hw_df, vec_logical))
mean(logical_num)

character_num = as.numeric(pull(hw_df, vec_character))
mean(character_num)

factor_num = as.numeric(pull(hw_df, vec_factor))
mean(factor_num)
```
When 




