---
title: "p8105_hw1_ei2262"
output: github_document
---

```{r, echo=FALSE, include = FALSE}
library(tidyverse)
```

# Problem 1

## Description of `penguins` Dataset

```{r, echo = FALSE}
data("penguins", package = "palmerpenguins")
#Viewing dataset (name/values)
penguins
```
The dataset `penguins` focuses on the following descriptive characteristics of penguins:

1. Type of species (Adélie, Chinstrap and Gentoo)
2. Island in Palmer Archipelago, Antarctica (Biscoe, Dream or Torgersen)
3. Bill length of penguin (in mm)
4. Bill depth  of penguin (in mm)
5. Flipper length of penguin (in mm)
6. Body mass of penguin (in g)
7. Sex of penguin (female or male)
8. Study year (2007, 2008, or 2009)

The dataset contains `r nrow(penguins)` rows and `r ncol(penguins)` columns. 
The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE),2)` mm.

## Scatterplot

The following is a scatterplot with `bill_length_mm` on the x-axis and
`flipper_length_mm`on the y-axis. The dataset is color coordinated based on the 
variable `species`.
```{r, echo = FALSE}
ggplot(data = penguins, aes(x=bill_length_mm,y=flipper_length_mm))+geom_point(aes(color=species))
```

The scatterplot was exported as a pdf to my project directory using `ggsave`. It was named "penguinsplot." 
```{r, echo = FALSE, include = FALSE}
ggsave("penguinsplot.pdf", units = "mm")
```

# Problem 2
```{r}
hw_df = tibble(
  vec_numeric = rnorm(n = 10),
  vec_logical = vec_numeric > 0,
  vec_character = c("d","s","h","o","m","e","w","o","r","k"),
  vec_factor = factor(c("high", "med","high","high","high","low","low","low","med","med"))
)

mean(pull(hw_df, vec_numeric))
mean(pull(hw_df, vec_logical))
mean(pull(hw_df, vec_character))
mean(pull(hw_df, vec_factor))
```
After creating the dataframe, I tried calculating the mean of each variable. I was able to calculate the mean for two variables:
*vec_numeric* and *vec_logical*. VWhen I calculated the mean for variables *vec_character* and *vec_factor*, I got an output of *NA*.

# Applying `as.numeric`
```{r, eval = FALSE, include = TRUE}
logical_num = as.numeric(pull(hw_df, vec_logical))
mean(logical_num)

character_num = as.numeric(pull(hw_df, vec_character))
mean(character_num)

factor_num = as.numeric(pull(hw_df, vec_factor))
mean(factor_num)
```
After applying the `as.numeric` function to the logical, character, and factor variables, I was able to calculate the mean of
the logical and factor variables. The character variable produces *NA* as the output.   
A logical vector has two possible levels, true or false. Therefore, when `as.numeric` is applied to a logical vector, it would be converted the observations to a 0 (false) or to a 1 (true). 0 and 1 are both numeric values and can be used to calculate a mean value.   
The factor vector created in the dataframe has 3 levels, high, medium, and low. When `as.numeric` is applied to it, the variable will convert each level to a corresponding number, such as 1 for high, 2 for medium, and 3 for low. Numbers 1 - 3 can then be used to calculate a mean value.  

This helps explain what happens when I am trying to take the mean and explains why I was unable to get a mean value before applying the `as.numeric` function to the vectors I created.





