---
title: "p8105_hw1_ei2262"
output: github_document
---

```{r, echo=FALSE, include = FALSE}
library(tidyverse)
```

# Problem 1

```{r, echo = FALSE}
data("penguins", package = "palmerpenguins")
#Viewing dataset (name/values)
penguins

#Number of rows 
nrow(penguins)
#Number of columns
ncol(penguins)

#Mean flipper length
mean(penguins$flipper_length_mm, na.rm = TRUE)
```
The dataset `penguins` focuses on the following descriptive characteristics of penguins:

1. Type of species
2. Island of origin
3. Bill length (in mm)
4. Bill depth (in mm)
5. Flipper length (in mm)
6. Body mass (in g)
7. Sex of penguin
8. Year

The dataset contains 344 rows and 8 columns. 
The mean flipper length is 200.92 mm.

The following is a scatterplot with `bill_length_mm` on the x-axis and
`flipper_length_mm`on the y-axis. The dataset is color coordinated based on the 
variable `species`.
```{r, echo = FALSE}
ggplot(data = penguins, aes(x=bill_length_mm,y=flipper_length_mm))+geom_point(aes(color=species))
```

```{r, echo = FALSE, include = FALSE}
ggsave("penguinsplot.pdf", units = "mm")
```


